### 2016-11-09 datamining/summery.tex
fixed geometry and added most baysian stuff I think

% This sets page margins to .5 inch if using letter paper, and to 1cm
% if using A4 paper. (This probably isn't strictly necessary.)
% If using another size paper, use default 1cm margins.
\ifthenelse{\lengthtest { \paperwidth = 11in}}
    { \geometry{top=.5in,left=.5in,right=.5in,bottom=.5in} }
    {\ifthenelse{ \lengthtest{ \paperwidth = 297mm}}
        {\geometry{top=0.5cm,left=0.5cm,right=0.5cm,bottom=0.5cm} }
        {\geometry{top=0.5cm,left=0.5cm,right=0.5cm,bottom=0.5cm} }
    }

@@

@@

\section{Freq item set mining}

>>>>

% This sets page margins to .5 inch if using letter paper, and to 1cm
% if using A4 paper. (This probably isn't strictly necessary.)
% If using another size paper, use default 1cm margins.
\geometry{top=0.2cm,left=0.2cm,right=0.2cm,bottom=0.2cm}

@@

\subsubsection{Iterative proportional fitting}
Is a converging algorithm
Start with a uniform table. Then set the row margins to the fitted counts
and distribute them proportionally. Then do the same for the columns, keep
doing this until the values converged.

\subsection{Decomosable graphical models}
Decomposable models are graphical models that have explicit formulas for the
maximum likelihood estimates.
This is a convenient property from a computational viewpoint.
If we only have to fit one model this is perhaps not so important,
but when we have little prior knowledge we typically have to search a
potentially large space of possible models.
Decomposable models are very easy to characterize by their independence graphs.
They have triangulated independence graphs: their independence graphs have no
chordless cycles of length greater than three. A cycle is called chordless
if no other than successive pairs of vertices in the cycle are adjacent. 
\subsection{Deviance and likelhood ratio test}
The deviance of a fitted model compares the log-likelihood of the fitted model
to the log-likelihood of the saturated model.
The larger the model deviance, the poorer the fit.
The likelihood of a model $M$ is
$L(\overset{\wedge}{P}(x);n(x))=\underset{x}{\prod}\overset{\wedge}{P}(x)^{n(x)}$
and the log likelihood of a model $M$ is
$\underset{x}{\sum} n(x)\log \overset{\wedge}{P}(x)$
$L^i=L(\overset{\wedge}{P}^{M_i})$ is the value of the log likelihood function
valutated at $\overset{\wedge}{P}^{M_i}$; the ML estimates of $P$ under $M_i$.
Let $M_0 \subseteq M_1$; ie $M_0$ can be obtained from $M_1$ by setting
additional $u$-terms to zero (more restrictions). The deviance between $M_0$ and
$M_1$ is $dev(M_0) - dev(M_1) = 2(L^1 - L^0)$.
For large $N$ $2(L^1-L^0) \approx M_0 \chi^2_v$ where the degrees of freedom $v$
is equal to the number of additional restrictions of $M_0$. 
We reject the null hypothesis that $M_0$ is the true model when:
$2(L^1-L^0)>\chi^2_{v;\alpha}$. 

@@

\subsection{Definition and notation}
A directed graph is a pair $G = (K, E)$, where $K$ is a set of vertices and $E$
is a set of edges with ordered pairs of vertices.
If there is an arrow from $i$ to $j$,
then we write this as $i \to j$, or equivalently as $(ij) \in E$.
We restrict attention to directed graphs with no directed cycles,
i.e. acyclic directed graphs (DAGs). If $i \to j$,
then $X_i$ is called a parent of $X_j$ ,
and $X_j$ is called a child of $X_i$ .
The set of coordinates of the parents of $X_j$ is denoted $pa(j)$, so
$X_{pa(j)}$ denotes the set of parents of $X_j$ .
If there is a directed path from $i$ to $j$,
then $X_i$ is called an ancestor of $X_j$ .
The set of coordinates of the ancestors of $X_j$ is denoted an $(j)$.
The absence of any directed cycles is equivalent to the existence of an
ordering of the nodes ${1, 2, \dots , k}$ such that $i \to j$ only when $i < j$.
In other words, there exists an ordering of the nodes such that arrows point
only from lower-numbered nodes to higher-numbered nodes.
Suppose that a priori knowledge tells us the variables can be labeled
$X_1 , X_2 , \dots , X_k$ such that $X_i$ is prior to $X_{i+1}$.
Corresponding to this ordering we can factorize the joint density of
$X_1 , X_2 , \dots , X_k$ as $P(X)=P(X_1)P(X_2|X_1)\dots P(X_k|X_{k-1},X_{k-2},\dots,X_1)$
In contructing a DAG, an arrow is drawn from $i$ to $j$, where $i<j$, unless
$P(X_j|X_{j-1}\dots,X_1)$ does not depend on $X_i$.
This is the key difference between DAGs and undirected graphs.
In both types of graph a missing edge between $X_i$ and $X_j$ is equivalent to
a conditional independence relation between $X_i$ and $X_j$.
In undirected graphs they are conditionally independent given all the remaining
variables, whereas in DAGs they are conditionally independent given all prior
variables. 
We can write the joint density:
$P(X)=\underset{i=1}{\overset{k}{\prod}}P(X_i|X_{pa(i)})$.
and the conditional indepencence as: $i \indep j|an({i,j})$
\subsection{Interpretation}
Given a DAG we construct the moral graph $G^m$ by marrying parents and deleting
directions, that is:
For each $i \in K$ we connect all vertices in $pa(i)$ with lines, we replace all
arrows in $E$ with lines.
\subsection{Maximum likelyhood estimation}
\subsubsection{ML estimation for Multinomal distribution}
We want to estimate the probabilities $p1 , p2 , \dots , p_J$
of getting outcomes $1, 2, \dots , J$ If in $n$ trials,
we observe n1 outcomes $1, n_2$ of $2, \dots, n_J$ of $J$,
then the obvious guess is to estimate $p_j , j = 1, \dots , J$, by $n_j /n$.
This is also the maximum likelihood estimate.
\subsection{ML estimation for bayesian nets}
The joint probability for $n$ independent observation in log likelyhood becomes
$L=\underset{i=1}{\overset{k}{\sum}}\underset{x_i,x_{pa(i)}}{\sum}n(x_i,x_{pa(i)})
\log p(x_i|x_{pa(i)})$
where $p$ stands for network parameters, $n(x_i,x_{pa(i)})$ for the number of
obsevations with $X_i=x_i$ and $X_{pa(i)}=x_{pa(i)}$.
If the parameters are not related this is just a bunch of independent
multinomalestimation problems. So we can say
$\overset{\wedge}{p}(x_i|x_{pa(i)})=\frac{n(x_i,x_{pa(i)})}{n(x_{pa(i))}}$,
which can be substituted in the $L$ formula.
\subsection{Estimation from incomplete data}
We paritition the complete data in an observed part $X_{obs}$ and a missing part
$X_{miss}$. For observation $j$ we write $X^{(j)}=(X_{obs}^{(j)},X_{mis}^{(j)})$.
We want to find those parameter values that maximize the probability of the
observed data.
This means that if some values are missing,
we have to obtain the marginal probability of the observed data by summing out
the missing data. For observation $j$,
the probability thus is:
$P(X^{j}_{obs})=\underset{X^{(j)}_{mis}}{\sum}P(X^{(j)})$
For all observations it becomes a product.
Direct maximization of the observed data likelihood is complicated because we get
a sum of parameters inside a log.
There is an algorithm iterative scheme to compute the ML estimate called
expectation maximization, it alternates between expectation step and maximization
steps. 
\subsection{Model selection with complete data}
To prevent overfitting we introduce a penalty. $L^M$ denotes the value of the
log-likelyhood function evaluated at $\overset{\wedge}{p}^M$; the ML estimates
of $p$ under model $M$. Akaikes information criterion would ive the follwoing
quality form model $M$: $AIC(M)=-2L^M+2dim(M)$ where $dim(M)$ is thenumber of
parameters of the model. which we divide by $-2$ to get higher value for higher
quality: $AIC(M)=L^M-dim(M)$. AIC gives a relativly low penalty for complex
models therefore Baysian infromation criterion is more popular:
$BIC(M)=L^M-\frac{\log n}{2} dim(M)$
So now we have a well-defined optimzation problem: Given training data, scoring
function (BIC), space of possible models (all DAG's) find a network that
maximaizes the score.
This is however NP hard therefore heuristics such as hillclimbing are used.
The score is decomposable, its a sum of terms where each term contains the
variables $i \cup pa(i)$. This means we only have to recalculate for variables
of which the parent set has changed.
\section{Freq paternset minig}
